---
title: "Integrating DRBayes with External Bayesian Software"
author: "Tomotaka Momozaki"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Integrating DRBayes with External Bayesian Software}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# 'cmdstanr' is not a CRAN package and CmdStan itself is not installed on CRAN
# check machines, so every chunk that compiles or runs a Stan model is guarded
# by `has_cmdstan`. When it is FALSE the vignette still knits; only the
# Stan-based output is omitted.
has_cmdstan <- requireNamespace("instantiate", quietly = TRUE) &&
  instantiate::stan_cmdstan_exists() &&
  requireNamespace("cmdstanr", quietly = TRUE)

# ggplot2 and dplyr live in Suggests, so they may be absent as well.
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
has_dplyr <- requireNamespace("dplyr", quietly = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
```

```{r cmdstan-note, echo = FALSE, results = "asis", eval = !has_cmdstan}
cat(
  "> **Note.** CmdStan (via the `cmdstanr` package) is not available in the\n",
  "> environment in which this vignette was built, so the code chunks that\n",
  "> compile and fit Stan models, and everything downstream of them, are shown\n",
  "> but not executed. The code is complete and runnable: install `cmdstanr`\n",
  "> and CmdStan and re-build the vignette to reproduce the numerical output.\n",
  sep = ""
)
```

# Introduction

The **DRBayes** package implements Bayesian doubly robust causal inference methods via posterior coupling. One of the key features of this package is its flexibility in working with posterior samples from various Bayesian software packages, including Stan, JAGS, and brms.

This vignette demonstrates how to integrate `drbayes_pc()` with external Bayesian software for both **continuous and binary outcomes**, allowing researchers to leverage their preferred modeling frameworks while benefiting from the doubly robust posterior coupling methodology.

## Key Features

- **Formula Interface**: Intuitive R formula syntax compatible with `lm()`, `glm()`, and other standard functions
- **Dual operation modes**: Internal sampling with built-in functions or external posterior samples
- **Multiple outcome types**: Support for both continuous and binary outcomes 
- **Flexible model families**: Works with Gaussian (linear) and binomial (logistic/probit) models 
- **Automatic detection**: Smart detection of outcome family and link functions based on data
- **Wide compatibility**: Works with Stan/cmdstanr, brms, JAGS, and other Bayesian software
- **Seamless integration**: Uses standard formula interface for consistent syntax

# Installation and Setup

```{r install, eval=FALSE}
# Install from GitHub (when available)
# devtools::install_github("t-momozaki/DRBayes")

# For Stan integration
# install.packages("cmdstanr", repos = c('https://stan-dev.r-universe.dev', getOption("repos")))
```

```{r load-packages}
# Load the DRBayes package
library(DRBayes)

# Load additional packages for external software integration
# library(cmdstanr)  # For Stan integration
# ggplot2 and dplyr are optional (Suggests); they are attached only if present
if (has_ggplot2) library(ggplot2)   # For visualization
if (has_dplyr) library(dplyr)       # For data manipulation and pipe operator (%>%)
```

# Data Setup

Let's generate both continuous and binary outcome examples that we'll use throughout this vignette:

```{r data-setup}
# Set seed for reproducibility
set.seed(1234)

# Generate synthetic data
n <- 500  # Sample size

# Create a comprehensive data frame
data <- data.frame(
  X1 = rnorm(n),
  X2 = rnorm(n),
  X3 = rnorm(n),
  X4 = rbinom(n, 1, 0.4)  # Binary covariate
)

# True treatment assignment mechanism
ps_logits <- 0.5 + 0.4*data$X1 + 0.3*data$X2 - 0.2*data$X3 + 0.5*data$X4
data$A <- rbinom(n, 1, plogis(ps_logits))

# Continuous outcome (ATE = 1.5)
data$Y_continuous <- 2 + 1.5*data$A + 0.8*data$X1 + 0.6*data$X2 + 0.4*data$X3 + 0.3*data$A*data$X1 + rnorm(n, 0, 1)

# Binary outcome; the true ATE in probability difference is computed below
linear_pred <- -0.5 + 0.8*data$A + 0.4*data$X1 + 0.3*data$X2 + 0.2*data$X3
data$Y_binary <- rbinom(n, 1, plogis(linear_pred))

# Calculate true ATE for binary outcome using counterfactual approach
prob_treated <- plogis(-0.5 + 0.8*1 + 0.4*data$X1 + 0.3*data$X2 + 0.2*data$X3)
prob_control <- plogis(-0.5 + 0.8*0 + 0.4*data$X1 + 0.3*data$X2 + 0.2*data$X3)
true_ate_binary <- mean(prob_treated - prob_control)

cat("Dataset created:\n")
cat("Sample size:", n, "\n")
cat("Treatment prevalence:", round(mean(data$A), 3), "\n")
cat("True ATE (continuous) =", 1.5, "\n")
cat("True ATE (binary, prob diff) =", round(true_ate_binary, 3), "\n")

# Display data structure
cat("\nData structure:\n")
str(data)
```

# Basic Usage: Internal Sampling

Before diving into external software integration, let's see how `drbayes_pc()` works with its built-in model functions for both outcome types:

```{r basic-usage}
# Continuous outcome using built-in models
result_internal_continuous <- drbayes_pc(
  outcome.formula = Y_continuous ~ A + X1 + X2 + X3 + A:X1,  # With interaction
  ps.formula = A ~ X1 + X2 + X3 + X4,                        # Include more predictors
  data = data,
  family = "gaussian",              # For continuous outcomes
  outcome.model = bayes_lm,             # Bayesian linear regression
  ps.model = bayes_logit,               # Bayesian logistic regression
  mc = 1200, bn = 300, thin = 2,
  outcome.priors = list(theta_prior = 1/100, sigma_prior = c(1, 1)),
  ps.priors = list(theta_prior = 1/100)
)

# Binary outcome using built-in models
result_internal_binary <- drbayes_pc(
  outcome.formula = Y_binary ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3 + X4,
  data = data,
  # family automatically detected as "binomial" since Y_binary is 0/1
  outcome.model = bayes_logit,          # Bayesian logistic regression
  ps.model = bayes_logit,               # Bayesian logistic regression
  mc = 1200, bn = 300, thin = 2,
  outcome.priors = list(theta_prior = 1/100),
  ps.priors = list(theta_prior = 1/100)
)

# Summarize results
cat("=== INTERNAL SAMPLING RESULTS ===\n")
cat("Continuous Outcome:\n")
cat("  Family used:", result_internal_continuous$family, "with", result_internal_continuous$link, "link\n")
cat("  ATE (PC): Mean =", round(mean(result_internal_continuous$pc), 3), 
    ", SD =", round(sd(result_internal_continuous$pc), 3), "\n")

cat("Binary Outcome:\n")
cat("  Family used:", result_internal_binary$family, "with", result_internal_binary$link, "link\n")
cat("  ATE (PC): Mean =", round(mean(result_internal_binary$pc), 3), 
    ", SD =", round(sd(result_internal_binary$pc), 3), "\n")
```

# External Software Integration

## Understanding Design Matrix Dimensions

When using external samples, it's crucial to understand how the formula interface creates design matrices:

```{r design-matrix-check}
# Check design matrix dimensions for our formulas
outcome_matrix_continuous <- model.matrix(Y_continuous ~ A + X1 + X2 + X3 + A:X1, data)
ps_matrix <- model.matrix(A ~ X1 + X2 + X3 + X4, data)
outcome_matrix_binary <- model.matrix(Y_binary ~ A + X1 + X2 + X3, data)

cat("Design Matrix Dimensions:\n")
cat("Continuous outcome model:", ncol(outcome_matrix_continuous), "parameters\n")
cat("Binary outcome model:", ncol(outcome_matrix_binary), "parameters\n")
cat("Propensity score model:", ncol(ps_matrix), "parameters\n")

# Show parameter names
cat("\nParameter names:\n")
cat("Continuous outcome:", colnames(outcome_matrix_continuous), "\n")
cat("Binary outcome:", colnames(outcome_matrix_binary), "\n")
cat("PS model:", colnames(ps_matrix), "\n")
```

## Stan/cmdstanr Integration

Stan is a popular platform for statistical modeling. Here's how to integrate Stan models with DRBayes for both continuous and binary outcomes:

### Step 1: Define Stan Models

```{r stan-models}
# Stan model for linear regression (continuous outcome)
stan_lm <- "
data {
  int<lower=0> N;              // number of observations
  int<lower=0> K;              // number of parameters
  matrix[N, K] X;              // design matrix (including intercept)
  vector[N] y;                 // response variable
}
parameters {
  vector[K] beta;              // regression coefficients
  real<lower=0> sigma;         // error standard deviation
}
model {
  // Priors
  beta ~ normal(0, sqrt(100)); // weakly informative priors
  sigma ~ inv_gamma(1, 1);     // prior for error variance
  
  // Likelihood
  y ~ normal(X * beta, sigma);
}
"

# Stan model for logistic regression (binary outcome & propensity score)
stan_logit <- "
data {
  int<lower=0> N;                     // number of observations
  int<lower=0> K;                     // number of parameters
  matrix[N, K] X;                     // design matrix (including intercept)
  array[N] int<lower=0, upper=1> y;   // binary response
}
parameters {
  vector[K] beta;              // regression coefficients
}
model {
  // Priors
  beta ~ normal(0, sqrt(100)); // weakly informative priors
  
  // Likelihood
  y ~ bernoulli_logit(X * beta);
}
"

# Stan model for probit regression (alternative for binary outcomes)
stan_probit <- "
data {
  int<lower=0> N;                     // number of observations
  int<lower=0> K;                     // number of parameters
  matrix[N, K] X;                     // design matrix (including intercept)
  array[N] int<lower=0, upper=1> y;   // binary response
}
parameters {
  vector[K] beta;              // regression coefficients
}
model {
  // Priors
  beta ~ normal(0, sqrt(100)); // weakly informative priors
  
  // Likelihood
  y ~ bernoulli(Phi(X * beta)); // Probit model
}
"
```

### Step 2: Compile and Fit Stan Models

```{r stan-fitting, eval = has_cmdstan}
# Compile models
mod_lm     <- cmdstanr::cmdstan_model(stan_file = cmdstanr::write_stan_file(stan_lm))
mod_logit  <- cmdstanr::cmdstan_model(stan_file = cmdstanr::write_stan_file(stan_logit))
mod_probit <- cmdstanr::cmdstan_model(stan_file = cmdstanr::write_stan_file(stan_probit))

cat("Fitting Stan models...\n")

# === CONTINUOUS OUTCOME MODEL ===
stan_data_continuous <- list(
  N = nrow(data),
  K = ncol(outcome_matrix_continuous),
  X = outcome_matrix_continuous,
  y = data$Y_continuous
)

fit_continuous_stan <- mod_lm$sample(
  data = stan_data_continuous,
  chains = 2,
  parallel_chains = 2,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

# === BINARY OUTCOME MODELS ===
stan_data_binary <- list(
  N = nrow(data),
  K = ncol(outcome_matrix_binary),
  X = outcome_matrix_binary,
  y = data$Y_binary
)

# Logistic regression
fit_binary_logit_stan <- mod_logit$sample(
  data = stan_data_binary,
  chains = 2,
  parallel_chains = 2,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

# Probit regression
fit_binary_probit_stan <- mod_probit$sample(
  data = stan_data_binary,
  chains = 2,
  parallel_chains = 2,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

# === PROPENSITY SCORE MODEL ===
stan_data_ps <- list(
  N = nrow(data),
  K = ncol(ps_matrix),
  X = ps_matrix,
  y = data$A
)

fit_ps_stan <- mod_logit$sample(
  data = stan_data_ps,
  chains = 2,
  parallel_chains = 2,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

cat("Stan model fitting completed!\n")
```

### Step 3: Extract Samples and Integrate with DRBayes

```{r stan-integration, eval = has_cmdstan}
# Extract samples from Stan fits
ps_samples_stan <- fit_ps_stan$draws("beta", format = "matrix")
outcome_samples_continuous <- fit_continuous_stan$draws("beta", format = "matrix")
outcome_samples_binary_logit <- fit_binary_logit_stan$draws("beta", format = "matrix")
outcome_samples_binary_probit <- fit_binary_probit_stan$draws("beta", format = "matrix")

# Verify sample dimensions
cat("Sample dimensions:\n")
cat("PS samples:", dim(ps_samples_stan), "\n")
cat("Continuous outcome samples:", dim(outcome_samples_continuous), "\n")
cat("Binary logit samples:", dim(outcome_samples_binary_logit), "\n")
cat("Binary probit samples:", dim(outcome_samples_binary_probit), "\n")

# === CONTINUOUS OUTCOME INTEGRATION ===
result_stan_continuous <- drbayes_pc(
  outcome.formula = Y_continuous ~ A + X1 + X2 + X3 + A:X1,
  ps.formula = A ~ X1 + X2 + X3 + X4,
  data = data,
  family = "gaussian",                    # Must specify for external samples
  outcome.samples = outcome_samples_continuous,
  ps.samples = ps_samples_stan,
  bn = 0, thin = 2  
)

# === BINARY OUTCOME INTEGRATION (LOGIT) ===
result_stan_binary_logit <- drbayes_pc(
  outcome.formula = Y_binary ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3 + X4,
  data = data,
  family = "binomial", 
  link = "logit",
  outcome.samples = outcome_samples_binary_logit,
  ps.samples = ps_samples_stan,
  bn = 0, thin = 2
)

# === BINARY OUTCOME INTEGRATION (PROBIT) ===
result_stan_binary_probit <- drbayes_pc(
  outcome.formula = Y_binary ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3 + X4,
  data = data,
  family = "binomial", 
  link = "probit",
  outcome.samples = outcome_samples_binary_probit,
  ps.samples = ps_samples_stan,
  bn = 0, thin = 2
)

# Summarize Stan integration results
cat("\n=== STAN INTEGRATION RESULTS ===\n")
cat("Continuous (Linear):\n")
cat("  ATE (PC): Mean =", round(mean(result_stan_continuous$pc), 3), 
    ", SD =", round(sd(result_stan_continuous$pc), 3), "\n")

cat("Binary (Logit):\n")
cat("  ATE (PC): Mean =", round(mean(result_stan_binary_logit$pc), 3), 
    ", SD =", round(sd(result_stan_binary_logit$pc), 3), "\n")

cat("Binary (Probit):\n")
cat("  ATE (PC): Mean =", round(mean(result_stan_binary_probit$pc), 3), 
    ", SD =", round(sd(result_stan_binary_probit$pc), 3), "\n")
```

# Comprehensive Comparison

```{r comprehensive-comparison, eval = has_cmdstan}
# Create a comprehensive comparison function
compare_all_results <- function() {
  
  results_list <- list(
    "Built-in Continuous" = result_internal_continuous,
    "Built-in Binary" = result_internal_binary,
    "Stan Continuous" = result_stan_continuous,
    "Stan Binary (Logit)" = result_stan_binary_logit,
    "Stan Binary (Probit)" = result_stan_binary_probit
  )
  
  outcome_types <- c("Continuous", "Binary", "Continuous", "Binary", "Binary")
  true_ates <- c(1.5, true_ate_binary, 1.5, true_ate_binary, true_ate_binary)
  
  comparison_df <- data.frame(
    Method = names(results_list),
    Outcome_Type = outcome_types,
    True_ATE = true_ates,
    PC_Mean = sapply(results_list, function(x) mean(x$pc)),
    PC_SD = sapply(results_list, function(x) sd(x$pc)),
    PC_CI_Lower = sapply(results_list, function(x) quantile(x$pc, 0.025)),
    PC_CI_Upper = sapply(results_list, function(x) quantile(x$pc, 0.975)),
    Family = sapply(results_list, function(x) x$family),
    Link = sapply(results_list, function(x) x$link)
  )
  
  # Calculate bias and coverage
  comparison_df$Bias <- comparison_df$PC_Mean - comparison_df$True_ATE
  comparison_df$Coverage <- mapply(function(true_val, ci_low, ci_high) {
    (true_val >= ci_low) & (true_val <= ci_high)
  }, comparison_df$True_ATE, comparison_df$PC_CI_Lower, comparison_df$PC_CI_Upper)
  
  # Round numeric columns
  numeric_cols <- c("True_ATE", "PC_Mean", "PC_SD", "PC_CI_Lower", "PC_CI_Upper", "Bias")
  comparison_df[, numeric_cols] <- round(comparison_df[, numeric_cols], 3)
  
  return(comparison_df)
}

comparison_table <- compare_all_results()
print(comparison_table)
```

```{r comparison-visualization, fig.width=14, fig.height=8, eval = has_cmdstan && has_ggplot2}
# Create comprehensive visualization
plot_data <- comparison_table
plot_data$Method_Clean <- gsub("Built-in ", "", plot_data$Method)
plot_data$Method_Clean <- gsub("Stan ", "", plot_data$Method_Clean)
plot_data$CI_Width <- plot_data$PC_CI_Upper - plot_data$PC_CI_Lower
plot_data$Software <- ifelse(grepl("Built-in", plot_data$Method), "Built-in", "Stan")

# Alternative using dplyr (if available)
if (require(dplyr, quietly = TRUE)) {
  plot_data <- comparison_table %>%
    mutate(
      Method_Clean = gsub("Built-in ", "", Method),
      Method_Clean = gsub("Stan ", "", Method_Clean),
      CI_Width = PC_CI_Upper - PC_CI_Lower,
      Software = ifelse(grepl("Built-in", Method), "Built-in", "Stan")
    )
}

# ATE comparison plot
p1 <- ggplot(plot_data, aes(x = Method_Clean, y = PC_Mean, color = Software)) +
  geom_point(size = 4, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = PC_CI_Lower, ymax = PC_CI_Upper), 
                width = 0.2, linewidth = 1, position = position_dodge(width = 0.3)) +
  geom_point(aes(y = True_ATE), shape = 4, size = 5, color = "red") +
  facet_wrap(~ Outcome_Type, scales = "free_y", ncol = 2) +
  labs(
    title = "ATE Estimates: Built-in vs External Software Integration",
    subtitle = "Points show posterior means, bars show 95% CIs, red X shows true ATE",
    x = "Method", 
    y = "ATE Estimate",
    color = "Software"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold")
  )

print(p1)
```

# Posterior Distribution Comparisons

```{r posterior-distributions, fig.width=14, fig.height=12, eval = has_cmdstan && has_ggplot2}
# Create data frame for all posterior distributions
create_posterior_data <- function() {
  data.frame(
    ATE = c(
      result_internal_continuous$pc,
      result_stan_continuous$pc,
      result_internal_binary$pc,
      result_stan_binary_logit$pc,
      result_stan_binary_probit$pc
    ),
    Method = rep(c("Built-in", "Stan", "Built-in", "Stan (Logit)", "Stan (Probit)"), 
                 times = c(
                   length(result_internal_continuous$pc),
                   length(result_stan_continuous$pc),
                   length(result_internal_binary$pc),
                   length(result_stan_binary_logit$pc),
                   length(result_stan_binary_probit$pc)
                 )),
    Outcome_Type = rep(c("Continuous", "Continuous", "Binary", "Binary", "Binary"),
                      times = c(
                        length(result_internal_continuous$pc),
                        length(result_stan_continuous$pc),
                        length(result_internal_binary$pc),
                        length(result_stan_binary_logit$pc),
                        length(result_stan_binary_probit$pc)
                      ))
  )
}

posterior_data <- create_posterior_data()

# Create density plots
p2 <- ggplot(posterior_data, aes(x = ATE, fill = Method)) +
  geom_density(alpha = 0.7) +
  geom_vline(data = data.frame(
    Outcome_Type = c("Continuous", "Binary"),
    true_ate = c(1.5, true_ate_binary)
  ), aes(xintercept = true_ate), linetype = "dashed", 
             color = "red", linewidth = 1) +
  facet_wrap(~ Outcome_Type, scales = "free", ncol = 1) +
  labs(
    title = "Posterior Distributions: Built-in vs External Software",
    subtitle = "Red line shows true ATE",
    x = "Average Treatment Effect",
    y = "Density",
    fill = "Method"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold")
  )

print(p2)
```

# Best Practices for External Integration

## Formula Interface Guidelines

### 1. Design Matrix Consistency
```r
# Always check design matrix dimensions before fitting external models
outcome_matrix <- model.matrix(your_outcome_formula, data)
ps_matrix <- model.matrix(your_ps_formula, data)

cat("Expected parameter counts:\n")
cat("Outcome model:", ncol(outcome_matrix), "\n")
cat("PS model:", ncol(ps_matrix), "\n")

# Your external software must produce samples with these exact dimensions
```

### 2. Family and Link Specification

`family` and `link` are auto-detected from the outcome, but the data cannot
say which link the external software used: a binary outcome could have come
from either a logit or a probit fit, and the two put the coefficients on
different scales. `ps.link` matters for the same reason. Name all three when
the samples come from elsewhere.

```r
result <- drbayes_pc(
  outcome.formula = Y ~ A + X1 + X2,
  ps.formula = A ~ X1 + X2 + X3,
  data = data,
  family = "binomial",       # auto-detected, but say it anyway
  link = "logit",            # the scale the outcome draws are on
  ps.link = "logit",         # the scale the propensity score draws are on
  outcome.samples = external_outcome_samples,
  ps.samples = external_ps_samples
)
```

### 3. Sample Dimension Validation
```r
# Create validation function
validate_external_samples <- function(outcome_samples, ps_samples, 
                                     outcome.formula, ps.formula, data) {
  outcome_matrix <- model.matrix(outcome.formula, data)
  ps_matrix <- model.matrix(ps.formula, data)
  
  errors <- c()
  
  if (ncol(outcome_samples) != ncol(outcome_matrix)) {
    errors <- c(errors, sprintf("Outcome samples: expected %d cols, got %d", 
                               ncol(outcome_matrix), ncol(outcome_samples)))
  }
  
  if (ncol(ps_samples) != ncol(ps_matrix)) {
    errors <- c(errors, sprintf("PS samples: expected %d cols, got %d", 
                               ncol(ps_matrix), ncol(ps_samples)))
  }
  
  if (nrow(outcome_samples) != nrow(ps_samples)) {
    errors <- c(errors, "Sample counts don't match between outcome and PS models")
  }
  
  if (length(errors) > 0) {
    stop(paste(errors, collapse = "; "))
  } else {
    cat("Sample dimensions validated successfully!\n")
  }
}

# Use validation
validate_external_samples(outcome_samples_continuous, ps_samples_stan,
                         Y_continuous ~ A + X1 + X2 + X3 + A:X1,
                         A ~ X1 + X2 + X3 + X4, data)
```

## Software-Specific Integration Tips

### Stan/cmdstanr
- Use `refresh = 0` for silent fitting in production
- Consider `parallel_chains = 4` for faster computation
- For binary outcomes, ensure correct likelihood specification

```r
# Best practices for Stan integration
# 1. Use consistent parameterization
# 2. Include intercepts in design matrices
# 3. Match DRBayes prior specifications when possible
# 4. Use 'format = "matrix"' for draws extraction

# Example Stan workflow
fit <- cmdstan_model(stan_file)$sample(
  data = list(N = nrow(data), K = ncol(design_matrix), 
              X = design_matrix, y = response),
  chains = 4,
  iter_sampling = 1000,
  refresh = 0
)

samples <- fit$draws("beta", format = "matrix")
```

## Troubleshooting Guide

```{r troubleshooting-examples, eval = has_cmdstan}
# 1. The design matrix and the samples disagree about how many coefficients
#    there are. This is the mistake the formula interface can catch, and it
#    names the two counts.
tryCatch({
  wrong_samples <- outcome_samples_continuous[, 1:3]  # the formula needs 6
  drbayes_pc(
    outcome.formula = Y_continuous ~ A + X1 + X2 + X3 + A:X1,
    ps.formula = A ~ X1 + X2 + X3 + X4,
    data = data,
    family = "gaussian",
    outcome.samples = wrong_samples,
    ps.samples = ps_samples_stan
  )
}, error = function(e) {
  cat("Expected error - dimension mismatch:\n")
  cat(conditionMessage(e), "\n\n")
})

# 2. The two sets of samples must have the same number of iterations, since a
#    tilted draw pairs one outcome draw with one propensity score draw.
tryCatch({
  drbayes_pc(
    outcome.formula = Y_continuous ~ A + X1 + X2 + X3 + A:X1,
    ps.formula = A ~ X1 + X2 + X3 + X4,
    data = data,
    family = "gaussian",
    outcome.samples = outcome_samples_continuous,
    ps.samples = ps_samples_stan[1:500, ],
    bn = 0
  )
}, error = function(e) {
  cat("Expected error - unequal iteration counts:\n")
  cat(conditionMessage(e), "\n\n")
})

# 3. What cannot be caught. Neither the formula nor the data records which
#    link the external software used, so probit draws handed over as logit
#    draws produce a number rather than an error. Name family, link and
#    ps.link yourself.
cat("Successful integration requires:\n")
cat("1. Correct sample dimensions\n")
cat("2. The same number of iterations in both sets of samples\n")
cat("3. family, link and ps.link matching the models that were fitted\n")
```

# Conclusion

The **DRBayes** package provides seamless integration with popular Bayesian software for both continuous and binary outcomes, allowing researchers to:

1. **Leverage existing expertise** with their preferred Bayesian software
2. **Handle multiple outcome types** with appropriate model families and link functions 
3. **Maintain modeling flexibility** while benefiting from doubly robust inference
4. **Compare results** across different software implementations and link functions
5. **Scale to complex models** that may be difficult to implement in built-in functions

The posterior coupling approach ensures that regardless of which software is used for initial model fitting, the final causal inference benefits from the doubly robust properties and uncertainty quantification that the method provides.

For more information about the theoretical foundations of the posterior coupling method, see:

> Orihara, S., Momozaki, T., & Sugasawa, S. (2025). Bayesian Doubly Robust Causal Inference via Posterior Coupling. arXiv preprint arXiv:2506.04868.

# Session Information

```{r session-info}
sessionInfo()
```
