---
title: "Integrating DRBayes with External Bayesian Software"
author: "Tomotaka Momozaki"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Integrating DRBayes with External Bayesian Software}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# Introduction

The **DRBayes** package implements Bayesian doubly robust causal inference methods via posterior coupling. One of the key features of this package is its flexibility in working with posterior samples from various Bayesian software packages, including Stan, JAGS, and brms.

This vignette demonstrates how to integrate `DRBayes.PC()` with external Bayesian software for both **continuous and binary outcomes**, allowing researchers to leverage their preferred modeling frameworks while benefiting from the doubly robust posterior coupling methodology.

## Key Features

- **Formula Interface**: Intuitive R formula syntax compatible with `lm()`, `glm()`, and other standard functions
- **Dual operation modes**: Internal sampling with built-in functions or external posterior samples
- **Multiple outcome types**: Support for both continuous and binary outcomes 
- **Flexible model families**: Works with Gaussian (linear) and binomial (logistic/probit) models 
- **Automatic detection**: Smart detection of outcome family and link functions based on data
- **Wide compatibility**: Works with Stan/cmdstanr, brms, JAGS, and other Bayesian software
- **Seamless integration**: Uses standard formula interface for consistent syntax

# Installation and Setup

```{r install, eval=FALSE}
# Install from GitHub (when available)
# devtools::install_github("t-momozaki/DRBayes")

# For Stan integration
# install.packages("cmdstanr", repos = c('https://stan-dev.r-universe.dev', getOption("repos")))
```

```{r load-packages}
# Load the DRBayes package
library(DRBayes)

# Load additional packages for external software integration
# library(cmdstanr)  # For Stan integration
library(ggplot2)   # For visualization
library(dplyr)     # For data manipulation and pipe operator (%>%)
```

# Data Setup

Let's generate both continuous and binary outcome examples that we'll use throughout this vignette:

```{r data-setup}
# Set seed for reproducibility
set.seed(1234)

# Generate synthetic data
n <- 2000  # Sample size

# Create a comprehensive data frame
data <- data.frame(
  X1 = rnorm(n),
  X2 = rnorm(n),
  X3 = rnorm(n),
  X4 = rbinom(n, 1, 0.4)  # Binary covariate
)

# True treatment assignment mechanism
ps_logits <- 0.5 + 0.4*data$X1 + 0.3*data$X2 - 0.2*data$X3 + 0.5*data$X4
data$A <- rbinom(n, 1, plogis(ps_logits))

# Continuous outcome (ATE = 1.5)
data$Y_continuous <- 2 + 1.5*data$A + 0.8*data$X1 + 0.6*data$X2 + 0.4*data$X3 + 0.3*data$A*data$X1 + rnorm(n, 0, 1)

# Binary outcome (ATE ≈ 0.18 in probability difference)
linear_pred <- -0.5 + 0.8*data$A + 0.4*data$X1 + 0.3*data$X2 + 0.2*data$X3
data$Y_binary <- rbinom(n, 1, plogis(linear_pred))

# Calculate true ATE for binary outcome using counterfactual approach
prob_treated <- plogis(-0.5 + 0.8*1 + 0.4*data$X1 + 0.3*data$X2 + 0.2*data$X3)
prob_control <- plogis(-0.5 + 0.8*0 + 0.4*data$X1 + 0.3*data$X2 + 0.2*data$X3)
true_ate_binary <- mean(prob_treated - prob_control)

cat("Dataset created:\n")
cat("Sample size:", n, "\n")
cat("Treatment prevalence:", round(mean(data$A), 3), "\n")
cat("True ATE (continuous) =", 1.5, "\n")
cat("True ATE (binary, prob diff) =", round(true_ate_binary, 3), "\n")

# Display data structure
cat("\nData structure:\n")
str(data)
```

# Basic Usage: Internal Sampling

Before diving into external software integration, let's see how `DRBayes.PC()` works with its built-in model functions for both outcome types:

```{r basic-usage}
# Continuous outcome using built-in models
result_internal_continuous <- DRBayes.PC(
  outcome.formula = Y_continuous ~ A + X1 + X2 + X3 + A:X1,  # With interaction
  ps.formula = A ~ X1 + X2 + X3 + X4,                        # Include more predictors
  data = data,
  family = "gaussian",              # For continuous outcomes
  outcome.model = B.LM,             # Bayesian linear regression
  ps.model = B.Logit,               # Bayesian logistic regression
  mc = 5000, bn = 1000, thin = 2,
  outcome.priors = list(theta_prior = 1/100, sigma_prior = c(1, 1)),
  ps.priors = list(theta_prior = 1/100)
)

# Binary outcome using built-in models
result_internal_binary <- DRBayes.PC(
  outcome.formula = Y_binary ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3 + X4,
  data = data,
  # family automatically detected as "binomial" since Y_binary ∈ {0,1}
  outcome.model = B.Logit,          # Bayesian logistic regression
  ps.model = B.Logit,               # Bayesian logistic regression
  mc = 5000, bn = 1000, thin = 2,
  outcome.priors = list(theta_prior = 1/100),
  ps.priors = list(theta_prior = 1/100)
)

# Summarize results
cat("=== INTERNAL SAMPLING RESULTS ===\n")
cat("Continuous Outcome:\n")
cat("  Family used:", result_internal_continuous$family, "with", result_internal_continuous$link, "link\n")
cat("  ATE (PC): Mean =", round(mean(result_internal_continuous$pc), 3), 
    ", SD =", round(sd(result_internal_continuous$pc), 3), "\n")

cat("Binary Outcome:\n")
cat("  Family used:", result_internal_binary$family, "with", result_internal_binary$link, "link\n")
cat("  ATE (PC): Mean =", round(mean(result_internal_binary$pc), 3), 
    ", SD =", round(sd(result_internal_binary$pc), 3), "\n")
```

# External Software Integration

## Understanding Design Matrix Dimensions

When using external samples, it's crucial to understand how the formula interface creates design matrices:

```{r design-matrix-check}
# Check design matrix dimensions for our formulas
outcome_matrix_continuous <- model.matrix(Y_continuous ~ A + X1 + X2 + X3 + A:X1, data)
ps_matrix <- model.matrix(A ~ X1 + X2 + X3 + X4, data)
outcome_matrix_binary <- model.matrix(Y_binary ~ A + X1 + X2 + X3, data)

cat("Design Matrix Dimensions:\n")
cat("Continuous outcome model:", ncol(outcome_matrix_continuous), "parameters\n")
cat("Binary outcome model:", ncol(outcome_matrix_binary), "parameters\n")
cat("Propensity score model:", ncol(ps_matrix), "parameters\n")

# Show parameter names
cat("\nParameter names:\n")
cat("Continuous outcome:", colnames(outcome_matrix_continuous), "\n")
cat("Binary outcome:", colnames(outcome_matrix_binary), "\n")
cat("PS model:", colnames(ps_matrix), "\n")
```

## Stan/cmdstanr Integration

Stan is a popular platform for statistical modeling. Here's how to integrate Stan models with DRBayes for both continuous and binary outcomes:

### Step 1: Define Stan Models

```{r stan-models}
# Stan model for linear regression (continuous outcome)
stan_lm <- "
data {
  int<lower=0> N;              // number of observations
  int<lower=0> K;              // number of parameters
  matrix[N, K] X;              // design matrix (including intercept)
  vector[N] y;                 // response variable
}
parameters {
  vector[K] beta;              // regression coefficients
  real<lower=0> sigma;         // error standard deviation
}
model {
  // Priors
  beta ~ normal(0, sqrt(100)); // weakly informative priors
  sigma ~ inv_gamma(1, 1);     // prior for error variance
  
  // Likelihood
  y ~ normal(X * beta, sigma);
}
"

# Stan model for logistic regression (binary outcome & propensity score)
stan_logit <- "
data {
  int<lower=0> N;                     // number of observations
  int<lower=0> K;                     // number of parameters
  matrix[N, K] X;                     // design matrix (including intercept)
  array[N] int<lower=0, upper=1> y;   // binary response
}
parameters {
  vector[K] beta;              // regression coefficients
}
model {
  // Priors
  beta ~ normal(0, sqrt(100)); // weakly informative priors
  
  // Likelihood
  y ~ bernoulli_logit(X * beta);
}
"

# Stan model for probit regression (alternative for binary outcomes)
stan_probit <- "
data {
  int<lower=0> N;                     // number of observations
  int<lower=0> K;                     // number of parameters
  matrix[N, K] X;                     // design matrix (including intercept)
  array[N] int<lower=0, upper=1> y;   // binary response
}
parameters {
  vector[K] beta;              // regression coefficients
}
model {
  // Priors
  beta ~ normal(0, sqrt(100)); // weakly informative priors
  
  // Likelihood
  y ~ bernoulli(Phi(X * beta)); // Probit model
}
"
```

### Step 2: Compile and Fit Stan Models

```{r stan-fitting}
# Compile models
mod_lm     <- cmdstanr::cmdstan_model(stan_file = cmdstanr::write_stan_file(stan_lm))
mod_logit  <- cmdstanr::cmdstan_model(stan_file = cmdstanr::write_stan_file(stan_logit))
mod_probit <- cmdstanr::cmdstan_model(stan_file = cmdstanr::write_stan_file(stan_probit))

cat("Fitting Stan models...\n")

# === CONTINUOUS OUTCOME MODEL ===
stan_data_continuous <- list(
  N = nrow(data),
  K = ncol(outcome_matrix_continuous),
  X = outcome_matrix_continuous,
  y = data$Y_continuous
)

fit_continuous_stan <- mod_lm$sample(
  data = stan_data_continuous,
  chains = 4,
  iter_sampling = 1000,
  iter_warmup = 250,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

# === BINARY OUTCOME MODELS ===
stan_data_binary <- list(
  N = nrow(data),
  K = ncol(outcome_matrix_binary),
  X = outcome_matrix_binary,
  y = data$Y_binary
)

# Logistic regression
fit_binary_logit_stan <- mod_logit$sample(
  data = stan_data_binary,
  chains = 4,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

# Probit regression
fit_binary_probit_stan <- mod_probit$sample(
  data = stan_data_binary,
  chains = 4,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

# === PROPENSITY SCORE MODEL ===
stan_data_ps <- list(
  N = nrow(data),
  K = ncol(ps_matrix),
  X = ps_matrix,
  y = data$A
)

fit_ps_stan <- mod_logit$sample(
  data = stan_data_ps,
  chains = 4,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

cat("Stan model fitting completed!\n")
```

### Step 3: Extract Samples and Integrate with DRBayes

```{r stan-integration}
# Extract samples from Stan fits
ps_samples_stan <- fit_ps_stan$draws("beta", format = "matrix")
outcome_samples_continuous <- fit_continuous_stan$draws("beta", format = "matrix")
outcome_samples_binary_logit <- fit_binary_logit_stan$draws("beta", format = "matrix")
outcome_samples_binary_probit <- fit_binary_probit_stan$draws("beta", format = "matrix")

# Verify sample dimensions
cat("Sample dimensions:\n")
cat("PS samples:", dim(ps_samples_stan), "\n")
cat("Continuous outcome samples:", dim(outcome_samples_continuous), "\n")
cat("Binary logit samples:", dim(outcome_samples_binary_logit), "\n")
cat("Binary probit samples:", dim(outcome_samples_binary_probit), "\n")

# === CONTINUOUS OUTCOME INTEGRATION ===
result_stan_continuous <- DRBayes.PC(
  outcome.formula = Y_continuous ~ A + X1 + X2 + X3 + A:X1,
  ps.formula = A ~ X1 + X2 + X3 + X4,
  data = data,
  family = "gaussian",                    # Must specify for external samples
  outcome.samples = outcome_samples_continuous,
  ps.samples = ps_samples_stan,
  bn = 0, thin = 2  
)

# === BINARY OUTCOME INTEGRATION (LOGIT) ===
result_stan_binary_logit <- DRBayes.PC(
  outcome.formula = Y_binary ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3 + X4,
  data = data,
  family = "binomial", 
  link = "logit",
  outcome.samples = outcome_samples_binary_logit,
  ps.samples = ps_samples_stan,
  bn = 0, thin = 2
)

# === BINARY OUTCOME INTEGRATION (PROBIT) ===
result_stan_binary_probit <- DRBayes.PC(
  outcome.formula = Y_binary ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3 + X4,
  data = data,
  family = "binomial", 
  link = "probit",
  outcome.samples = outcome_samples_binary_probit,
  ps.samples = ps_samples_stan,
  bn = 0, thin = 2
)

# Summarize Stan integration results
cat("\n=== STAN INTEGRATION RESULTS ===\n")
cat("Continuous (Linear):\n")
cat("  ATE (PC): Mean =", round(mean(result_stan_continuous$pc), 3), 
    ", SD =", round(sd(result_stan_continuous$pc), 3), "\n")

cat("Binary (Logit):\n")
cat("  ATE (PC): Mean =", round(mean(result_stan_binary_logit$pc), 3), 
    ", SD =", round(sd(result_stan_binary_logit$pc), 3), "\n")

cat("Binary (Probit):\n")
cat("  ATE (PC): Mean =", round(mean(result_stan_binary_probit$pc), 3), 
    ", SD =", round(sd(result_stan_binary_probit$pc), 3), "\n")
```

# Advanced Formula Usage with External Software

## Complex Transformations and Interactions

```{r advanced-formulas-stan}
# Create data with more complex relationships
set.seed(9876)
n_complex <- 300

data_complex <- data.frame(
  X1 = rnorm(n_complex, 2, 0.5),  # For log transformation
  X2 = rnorm(n_complex),
  X3 = rnorm(n_complex),
  X4 = runif(n_complex, 0, 5)   # For polynomial
)

# Treatment
data_complex$A <- rbinom(n_complex, 1, plogis(0.2 + 0.3*data_complex$X1 + 0.2*data_complex$X2^2))

# Outcome with complex relationships
data_complex$Y <- 1.5 + 1.2*data_complex$A + 
                  0.4*log(data_complex$X1) + 
                  0.3*data_complex$X2^2 + 
                  0.2*data_complex$X3 + 
                  0.1*data_complex$X4^2 +
                  0.5*data_complex$A*data_complex$X1 +
                  rnorm(n_complex, 0, 0.8)

# Check the complex design matrix
complex_outcome_matrix <- model.matrix(Y ~ A + log(X1) + I(X2^2) + X3 + I(X4^2) + A:X1, data_complex)
complex_ps_matrix <- model.matrix(A ~ X1 + I(X2^2), data_complex)

cat("Complex formula design matrices:\n")
cat("Outcome parameters:", ncol(complex_outcome_matrix), "\n")
cat("PS parameters:", ncol(complex_ps_matrix), "\n")
cat("Parameter names:", colnames(complex_outcome_matrix), "\n")

# Fit complex model with Stan
stan_data_complex <- list(
  N = nrow(data_complex),
  K = ncol(complex_outcome_matrix),
  X = complex_outcome_matrix,
  y = data_complex$Y
)

fit_complex_stan <- mod_lm$sample(
  data = stan_data_complex,
  chains = 4,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

# PS model for complex data
stan_data_complex_ps <- list(
  N = nrow(data_complex),
  K = ncol(complex_ps_matrix),
  X = complex_ps_matrix,
  y = data_complex$A
)

fit_complex_ps_stan <- mod_logit$sample(
  data = stan_data_complex_ps,
  chains = 4,
  iter_sampling = 1000,
  iter_warmup = 1000,
  refresh = 0,
  show_messages = FALSE, 
  show_exceptions = FALSE
)

# Extract samples and use with DRBayes
complex_outcome_samples <- fit_complex_stan$draws("beta", format = "matrix")
complex_ps_samples <- fit_complex_ps_stan$draws("beta", format = "matrix")

result_complex_stan <- DRBayes.PC(
  outcome.formula = Y ~ A + log(X1) + I(X2^2) + X3 + I(X4^2) + A:X1,
  ps.formula = A ~ X1 + I(X2^2),
  data = data_complex,
  family = "gaussian",
  outcome.samples = complex_outcome_samples,
  ps.samples = complex_ps_samples,
  bn = 0, thin = 2
)

cat("Complex formula with Stan:\n")
cat("ATE estimate:", round(mean(result_complex_stan$pc), 3), "\n")
```


# Comprehensive Comparison

```{r comprehensive-comparison}
# Create a comprehensive comparison function
compare_all_results <- function() {
  
  results_list <- list(
    "Built-in Continuous" = result_internal_continuous,
    "Built-in Binary" = result_internal_binary,
    "Stan Continuous" = result_stan_continuous,
    "Stan Binary (Logit)" = result_stan_binary_logit,
    "Stan Binary (Probit)" = result_stan_binary_probit,
    "Stan Complex" = result_complex_stan
  )
  
  outcome_types <- c("Continuous", "Binary", "Continuous", "Binary", "Binary", "Continuous")
  true_ates <- c(1.5, true_ate_binary, 1.5, true_ate_binary, true_ate_binary, 1.2) # Approximate for complex
  
  comparison_df <- data.frame(
    Method = names(results_list),
    Outcome_Type = outcome_types,
    True_ATE = true_ates,
    PC_Mean = sapply(results_list, function(x) mean(x$pc)),
    PC_SD = sapply(results_list, function(x) sd(x$pc)),
    PC_CI_Lower = sapply(results_list, function(x) quantile(x$pc, 0.025)),
    PC_CI_Upper = sapply(results_list, function(x) quantile(x$pc, 0.975)),
    Family = sapply(results_list, function(x) x$family),
    Link = sapply(results_list, function(x) x$link)
  )
  
  # Calculate bias and coverage
  comparison_df$Bias <- comparison_df$PC_Mean - comparison_df$True_ATE
  comparison_df$Coverage <- mapply(function(true_val, ci_low, ci_high) {
    (true_val >= ci_low) & (true_val <= ci_high)
  }, comparison_df$True_ATE, comparison_df$PC_CI_Lower, comparison_df$PC_CI_Upper)
  
  # Round numeric columns
  numeric_cols <- c("True_ATE", "PC_Mean", "PC_SD", "PC_CI_Lower", "PC_CI_Upper", "Bias")
  comparison_df[, numeric_cols] <- round(comparison_df[, numeric_cols], 3)
  
  return(comparison_df)
}

comparison_table <- compare_all_results()
print(comparison_table)
```

```{r comparison-visualization, fig.width=14, fig.height=10}
# Create comprehensive visualization
plot_data <- comparison_table %>%
  mutate(
    Method_Clean = gsub("Built-in ", "", Method),
    Method_Clean = gsub("Stan ", "", Method_Clean),
    CI_Width = PC_CI_Upper - PC_CI_Lower,
    Software = ifelse(grepl("Built-in", Method), "Built-in", "Stan")
  )

# ATE comparison plot
p1 <- ggplot(plot_data, aes(x = Method_Clean, y = PC_Mean, color = Software, shape = Outcome_Type)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = PC_CI_Lower, ymax = PC_CI_Upper), 
                width = 0.2, linewidth = 1) +
  geom_point(aes(y = True_ATE), shape = 4, size = 5, color = "red") +
  facet_wrap(~ Outcome_Type, scales = "free_y", ncol = 2) +
  labs(
    title = "ATE Estimates: Built-in vs External Software Integration",
    subtitle = "Points show posterior means, bars show 95% CIs, red X shows true ATE",
    x = "Method", 
    y = "ATE Estimate",
    color = "Software",
    shape = "Outcome Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold")
  )

print(p1)

# Bias comparison
p2 <- ggplot(plot_data, aes(x = Method_Clean, y = Bias, fill = Software)) +
  geom_col(position = "dodge", alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  facet_wrap(~ Outcome_Type, scales = "free_y") +
  labs(
    title = "Bias Comparison Across Methods",
    x = "Method",
    y = "Bias (Estimate - True ATE)",
    fill = "Software"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

print(p2)
```

# Posterior Distribution Comparisons

```{r posterior-distributions, fig.width=14, fig.height=12}
# Create data frame for all posterior distributions
create_posterior_data <- function() {
  data.frame(
    ATE = c(
      result_internal_continuous$pc,
      result_stan_continuous$pc,
      result_internal_binary$pc,
      result_stan_binary_logit$pc,
      result_stan_binary_probit$pc
    ),
    Method = rep(c("Built-in", "Stan", "Built-in", "Stan (Logit)", "Stan (Probit)"), 
                 times = c(
                   length(result_internal_continuous$pc),
                   length(result_stan_continuous$pc),
                   length(result_internal_binary$pc),
                   length(result_stan_binary_logit$pc),
                   length(result_stan_binary_probit$pc)
                 )),
    Outcome_Type = rep(c("Continuous", "Continuous", "Binary", "Binary", "Binary"),
                      times = c(
                        length(result_internal_continuous$pc),
                        length(result_stan_continuous$pc),
                        length(result_internal_binary$pc),
                        length(result_stan_binary_logit$pc),
                        length(result_stan_binary_probit$pc)
                      ))
  )
}

posterior_data <- create_posterior_data()

# Create density plots
p3 <- ggplot(posterior_data, aes(x = ATE, fill = Method)) +
  geom_density(alpha = 0.7) +
  geom_vline(data = data.frame(
    Outcome_Type = c("Continuous", "Binary"),
    true_ate = c(1.5, true_ate_binary)
  ), aes(xintercept = true_ate), linetype = "dashed", 
             color = "red", linewidth = 1) +
  facet_wrap(~ Outcome_Type, scales = "free", ncol = 1) +
  labs(
    title = "Posterior Distributions: Built-in vs External Software",
    subtitle = "Red line shows true ATE",
    x = "Average Treatment Effect",
    y = "Density",
    fill = "Method"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold")
  )

print(p3)
```

# Best Practices for External Integration

## Formula Interface Guidelines

### 1. Design Matrix Consistency
```r
# Always check design matrix dimensions before fitting external models
outcome_matrix <- model.matrix(your_outcome_formula, data)
ps_matrix <- model.matrix(your_ps_formula, data)

cat("Expected parameter counts:\n")
cat("Outcome model:", ncol(outcome_matrix), "\n")
cat("PS model:", ncol(ps_matrix), "\n")

# Your external software must produce samples with these exact dimensions
```

### 2. Family and Link Specification
```r
# For external samples, always specify family and link explicitly
result <- DRBayes.PC(
  outcome.formula = Y ~ A + X1 + X2,
  ps.formula = A ~ X1 + X2 + X3,
  data = data,
  family = "binomial",        # Required for external samples
  link = "logit",            # Required for external samples
  outcome.samples = external_outcome_samples,
  ps.samples = external_ps_samples
)
```

### 3. Sample Dimension Validation
```r
# Create validation function
validate_external_samples <- function(outcome_samples, ps_samples, 
                                     outcome.formula, ps.formula, data) {
  outcome_matrix <- model.matrix(outcome.formula, data)
  ps_matrix <- model.matrix(ps.formula, data)
  
  errors <- c()
  
  if (ncol(outcome_samples) != ncol(outcome_matrix)) {
    errors <- c(errors, sprintf("Outcome samples: expected %d cols, got %d", 
                               ncol(outcome_matrix), ncol(outcome_samples)))
  }
  
  if (ncol(ps_samples) != ncol(ps_matrix)) {
    errors <- c(errors, sprintf("PS samples: expected %d cols, got %d", 
                               ncol(ps_matrix), ncol(ps_samples)))
  }
  
  if (nrow(outcome_samples) != nrow(ps_samples)) {
    errors <- c(errors, "Sample counts don't match between outcome and PS models")
  }
  
  if (length(errors) > 0) {
    stop(paste(errors, collapse = "; "))
  } else {
    cat("Sample dimensions validated successfully!\n")
  }
}

# Use validation
validate_external_samples(outcome_samples_continuous, ps_samples_stan,
                         Y_continuous ~ A + X1 + X2 + X3 + A:X1,
                         A ~ X1 + X2 + X3 + X4, data)
```

## Software-Specific Integration Tips

### Stan/cmdstanr
- Use `refresh = 0` for silent fitting in production
- Consider `parallel_chains = 4` for faster computation
- For binary outcomes, ensure correct likelihood specification

```r
# Best practices for Stan integration
# 1. Use consistent parameterization
# 2. Include intercepts in design matrices
# 3. Match DRBayes prior specifications when possible
# 4. Use 'format = "matrix"' for draws extraction

# Example Stan workflow
fit <- cmdstan_model(stan_file)$sample(
  data = list(N = nrow(data), K = ncol(design_matrix), 
              X = design_matrix, y = response),
  chains = 4,
  iter_sampling = 1000,
  refresh = 0
)

samples <- fit$draws("beta", format = "matrix")
```

## Troubleshooting Guide

```{r troubleshooting-examples}
# Common issues and solutions

# 1. Dimension mismatch error
tryCatch({
  # This will fail - wrong number of columns
  wrong_samples <- outcome_samples_continuous[, 1:3]  # Only 3 columns instead of required
  DRBayes.PC(
    outcome.formula = Y_continuous ~ A + X1 + X2 + X3 + A:X1,
    ps.formula = A ~ X1 + X2 + X3 + X4,
    data = data,
    family = "gaussian",
    outcome.samples = wrong_samples,
    ps.samples = ps_samples_stan
  )
}, error = function(e) {
  cat("Expected error - dimension mismatch:\n")
  cat(e$message, "\n\n")
})

# 2. Missing family specification
tryCatch({
  DRBayes.PC(
    outcome.formula = Y_continuous ~ A + X1 + X2 + X3,
    ps.formula = A ~ X1 + X2 + X3 + X4,
    data = data,
    # Missing family specification for external samples
    outcome.samples = outcome_samples_continuous[, 1:5], # Adjust to right size
    ps.samples = ps_samples_stan
  )
}, error = function(e) {
  cat("Expected error - missing family specification:\n")
  cat(e$message, "\n\n")
})

# 3. Successful validation
cat("Successful integration requires:\n")
cat("1. Correct sample dimensions\n")
cat("2. Explicit family and link specification\n")
cat("3. Consistent formula specifications\n")
```

# Conclusion

The **DRBayes** package provides seamless integration with popular Bayesian software for both continuous and binary outcomes, allowing researchers to:

1. **Leverage existing expertise** with their preferred Bayesian software
2. **Handle multiple outcome types** with appropriate model families and link functions 
3. **Maintain modeling flexibility** while benefiting from doubly robust inference
4. **Compare results** across different software implementations and link functions
5. **Scale to complex models** that may be difficult to implement in built-in functions

The posterior coupling approach ensures that regardless of which software is used for initial model fitting, the final causal inference benefits from the doubly robust properties and uncertainty quantification that the method provides.

For more information about the theoretical foundations of the posterior coupling method, see:

> Orihara, S., Momozaki, T., & Sugasawa, S. (2025). Bayesian Doubly Robust Causal Inference via Posterior Coupling. arXiv preprint arXiv:2506.04868.

# Session Information

```{r session-info}
sessionInfo()
```
