---
title: "Getting Started with DRBayes"
author: "Tomotaka Momozaki"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Getting Started with DRBayes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# Introduction

The **DRBayes** package implements Bayesian doubly robust causal inference methods via posterior coupling. This approach combines the robustness of doubly robust estimation with the uncertainty quantification benefits of Bayesian inference, providing a principled framework for causal effect estimation.

## Key Features

- **Formula Interface**: Intuitive R formula syntax compatible with `lm()`, `glm()`, and other standard functions
- **Doubly robust estimation**: Consistent estimation even when either the outcome model or propensity score model is misspecified
- **Bayesian uncertainty quantification**: Full posterior distributions for treatment effects
- **Posterior coupling**: Novel methodology that avoids the feedback problem in traditional doubly robust methods
- **Multiple outcome types**: Support for both continuous (linear regression) and binary outcomes (logistic/probit regression) 
- **Automatic detection**: Smart detection of outcome family and link functions based on data
- **Flexible modeling**: Built-in functions with standard and horseshoe priors plus support for external Bayesian software

## Installation

```{r install, eval=FALSE}
# Install from GitHub (when available)
# devtools::install_github("t-momozaki/DRBayes")
```

```{r load-packages}
# Load the DRBayes package
library(DRBayes)

# Load additional packages for visualization
library(ggplot2)
```

# Basic Concepts

## Doubly Robust Estimation

Doubly robust (DR) estimation combines two models:

1. **Outcome model**: $E[Y|A,X] = \mu(A,X)$ - models the relationship between outcome and covariates
2. **Propensity score model**: $P(A=1|X) = \pi(X)$ - models the treatment assignment mechanism

The key property is that DR estimators remain consistent if **either** model is correctly specified, providing robustness against model misspecification.

## Outcome Types and Model Families 

DRBayes supports different outcome types through the `family` and `link` parameters:

### Continuous Outcomes (Gaussian Family)
- **Family**: `"gaussian"`
- **Link**: `"identity"` (default)
- **Model**: Linear regression
- **ATE Interpretation**: Mean difference $E[Y|A=1,X] - E[Y|A=0,X]$

### Binary Outcomes (Binomial Family)
- **Family**: `"binomial"`
- **Link**: `"logit"` (default) or `"probit"`
- **Model**: Logistic or probit regression
- **ATE Interpretation**: Probability difference $P(Y=1|A=1,X) - P(Y=1|A=0,X)$

## Posterior Coupling Approach

Traditional DR methods can suffer from the "feedback problem" where the two models influence each other in undesirable ways. The posterior coupling approach in DRBayes:

- Generates posterior samples from both models separately
- Uses sequential Monte Carlo to enforce the DR moment condition
- Produces a coupled posterior that satisfies the DR property

## Formula Interface

The formula interface in DRBayes provides several advantages:

- **Intuitive syntax**: Similar to `lm()` and `glm()`
- **Automatic handling**: Design matrices, factor variables, interactions
- **Missing data support**: Built-in handling of missing values
- **Complex transformations**: Support for `I()`, `log()`, `poly()`, etc.

# Example 1: Continuous Outcomes

## Data Setup

Let's start with a continuous outcome example using simulated data:

```{r data-generation-continuous}
# Set seed for reproducibility
set.seed(1234)

# Sample size
n <- 2000

# Create a data frame with all variables
data <- data.frame(
  X1 = rnorm(n),
  X2 = rnorm(n),
  X3 = rbinom(n, 1, 0.5),
  X4 = rnorm(n)  # Additional covariate
)

# True propensity score (treatment assignment mechanism)
ps_true <- plogis(0.2 + 0.5*data$X1 - 0.3*data$X2 + 0.4*data$X3)
data$A <- rbinom(n, 1, ps_true)

# True outcome model with treatment effect (CONTINUOUS)
# ATE = 1.5 (main effect) + interaction effects
data$Y <- 1 + 1.5*data$A + 0.8*data$X1 + 0.6*data$X2 + 0.4*data$X3 + 0.3*data$A*data$X1 + rnorm(n, 0, 1)

# Calculate true ATE (accounting for interactions)
# E[Y|A=1,X] - E[Y|A=0,X] = 1.5 + 0.3*E[X1] = 1.5 (since E[X1] = 0)
true_ate_continuous <- 1.5 

cat("True ATE (continuous):", round(true_ate_continuous, 3), "\n")
cat("Sample statistics:\n")
cat("  Treatment group mean:", round(mean(data$Y[data$A==1]), 3), "\n")
cat("  Control group mean:", round(mean(data$Y[data$A==0]), 3), "\n")
cat("  Observed difference:", round(mean(data$Y[data$A==1]) - mean(data$Y[data$A==0]), 3), "\n")
cat("  Sample size:", nrow(data), "\n")
cat("  Treatment prevalence:", round(mean(data$A), 3), "\n")
```

## Fitting the Model with Formula Interface

```{r model-fitting-continuous}
# Fit the model using the formula interface
result_continuous <- DRBayes.PC(
  outcome.formula = Y ~ A + X1 + X2 + X3 + A:X1,  # Outcome model with interaction
  ps.formula = A ~ X1 + X2 + X3,                  # Propensity score model
  data = data,
  family = "gaussian",                            # For continuous outcomes
  link = "identity",                              # Identity link (default)
  outcome.model = B.LM,                           # Bayesian linear regression
  ps.model = B.Logit,                             # Bayesian logistic regression
  mc = 5000,                                      # MCMC iterations
  bn = 1000,                                      # Burn-in
  thin = 2,                                       # Thinning
  outcome.priors = list(
    theta_prior = 1/100,                          # Prior precision for regression coefficients
    sigma_prior = c(1, 1)                         # Inverse-gamma prior for error variance
  ),
  ps.priors = list(
    theta_prior = 1/100                           # Prior precision for logistic regression
  )
)

cat("Continuous outcome model fitting completed!\n")
cat("Number of posterior samples:", length(result_continuous$g.comp), "\n")
cat("Model family:", result_continuous$family, "\n")
cat("Link function:", result_continuous$link, "\n")
```

## Automatic Family Detection

```{r auto-detection-continuous}
# DRBayes can automatically detect the outcome type
result_auto_continuous <- DRBayes.PC(
  outcome.formula = Y ~ A + X1 + X2 + X3 + A:X1,
  ps.formula = A ~ X1 + X2 + X3,
  data = data,
  # family and link not specified - will be auto-detected
  outcome.model = B.LM, 
  ps.model = B.Logit,
  mc = 1000, bn = 200, thin = 2
)

cat("Auto-detected family:", result_auto_continuous$family, "\n")
cat("Auto-detected link:", result_auto_continuous$link, "\n")
cat("Auto-detected ATE:", round(mean(result_auto_continuous$pc), 3), "\n")
```

# Example 2: Binary Outcomes 

## Data Setup (Binary)

Now let's create a binary outcome example:

```{r data-generation-binary}
# Create new data for binary outcome
set.seed(5678)
n <- 2000

data_binary <- data.frame(
  X1 = rnorm(n),
  X2 = rnorm(n),
  X3 = rbinom(n, 1, 0.5)
)

# Treatment assignment (same mechanism)
ps_true_bin <- plogis(0.2 + 0.5*data_binary$X1 - 0.3*data_binary$X2 + 0.4*data_binary$X3)
data_binary$A <- rbinom(n, 1, ps_true_bin)

# Binary outcome using logistic model
linear_predictor <- -0.5 + 0.8*data_binary$A + 0.4*data_binary$X1 + 0.3*data_binary$X2 + 0.2*data_binary$X3
prob_success <- plogis(linear_predictor)
data_binary$Y <- rbinom(n, 1, prob_success)

# Calculate true ATE for binary outcome (probability difference)
# Create counterfactual outcomes
prob_treated <- plogis(-0.5 + 0.8*1 + 0.4*data_binary$X1 + 0.3*data_binary$X2 + 0.2*data_binary$X3)
prob_control <- plogis(-0.5 + 0.8*0 + 0.4*data_binary$X1 + 0.3*data_binary$X2 + 0.2*data_binary$X3)
true_ate_binary <- mean(prob_treated - prob_control)

cat("True ATE (binary, probability difference):", round(true_ate_binary, 3), "\n")
cat("Sample statistics:\n")
cat("  Treatment group success rate:", round(mean(data_binary$Y[data_binary$A==1]), 3), "\n")
cat("  Control group success rate:", round(mean(data_binary$Y[data_binary$A==0]), 3), "\n")
cat("  Observed difference:", round(mean(data_binary$Y[data_binary$A==1]) - mean(data_binary$Y[data_binary$A==0]), 3), "\n")
```

## Fitting Binary Outcome Models

```{r model-fitting-binary}
# Fit logistic model - will be auto-detected
result_binary_auto <- DRBayes.PC(
  outcome.formula = Y ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3,
  data = data_binary,
  # Auto-detection: all Y values are 0/1 → binomial family with logit link
  outcome.model = B.Logit,
  ps.model = B.Logit,
  mc = 5000, bn = 1000, thin = 2
)

cat("Auto-detected family:", result_binary_auto$family, "\n")
cat("Auto-detected link:", result_binary_auto$link, "\n")

# Fit with explicit specification
result_binary_logit <- DRBayes.PC(
  outcome.formula = Y ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3,
  data = data_binary,
  family = "binomial",
  link = "logit",
  outcome.model = B.Logit,
  ps.model = B.Logit,
  mc = 5000, bn = 1000, thin = 2
)

# Fit with probit link
result_binary_probit <- DRBayes.PC(
  outcome.formula = Y ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3,
  data = data_binary,
  family = "binomial",
  link = "probit",
  outcome.model = B.Probit,
  ps.model = B.Logit,
  mc = 5000, bn = 1000, thin = 2
)

cat("Binary outcome models completed!\n")
```

# Advanced Formula Usage

## Complex Transformations and Interactions

```{r advanced-formulas}
# Generate more complex data
set.seed(9999)
n <- 500

data_advanced <- data.frame(
  X1 = rnorm(n, 2, 0.5),     # Mean-shifted for log transformation
  X2 = rnorm(n),
  X3 = rnorm(n),
  X4 = runif(n, 0, 5)      # Uniform for polynomial
)

# Treatment assignment
ps_adv <- plogis(0.1 + 0.3*data_advanced$X1 + 0.2*data_advanced$X2^2 + 0.1*data_advanced$X3*data_advanced$X4)
data_advanced$A <- rbinom(n, 1, ps_adv)

# Complex outcome with non-linear relationships
data_advanced$Y <- 2 + 1.2*data_advanced$A + 
                   0.5*log(data_advanced$X1) + 
                   0.3*data_advanced$X2^2 + 
                   0.4*data_advanced$X3 + 
                   0.2*data_advanced$X4^2 + 
                   0.6*data_advanced$A*data_advanced$X1 + 
                   rnorm(n, 0, 0.8)

# Fit with complex formulas
result_advanced <- DRBayes.PC(
  outcome.formula = Y ~ A + log(X1) + I(X2^2) + X3 + poly(X4, 2) + A:X1,
  ps.formula = A ~ X1 + X2 + X3 + I(X2^2) + X3:X4,
  data = data_advanced,
  family = "gaussian",
  outcome.model = B.LM,
  ps.model = B.Logit,
  mc = 2000, bn = 500, thin = 2
)

cat("Advanced formula model completed!\n")
cat("ATE estimate:", round(mean(result_advanced$pc), 3), "\n")
```

## Missing Data Handling

```{r missing-data}
# Create data with missing values
data_missing <- data
missing_indices_x1 <- sample(1:length(data_missing$Y), 30)
missing_indices_x2 <- sample(1:length(data_missing$Y), 25)
data_missing$X1[missing_indices_x1] <- NA
data_missing$X2[missing_indices_x2] <- NA

cat("Original sample size:", length(data_missing$Y), "\n")
cat("Missing values in X1:", length(missing_indices_x1), "\n")
cat("Missing values in X2:", length(missing_indices_x2), "\n")

# Fit model with missing data using na.omit
result_missing <- DRBayes.PC(
  outcome.formula = Y ~ A + X1 + X2 + X3,
  ps.formula = A ~ X1 + X2 + X3,
  data = data_missing,
  na.action = "na.omit",
  outcome.model = B.LM,
  ps.model = B.Logit,
  mc = 2000, bn = 500, thin = 2
)

cat("After handling missing data:\n")
cat("Analysis sample size:", result_missing$data_info$n_observations, "\n")
cat("Excluded observations:", result_missing$data_info$missing_observations, "\n")
cat("ATE with complete cases:", round(mean(result_missing$pc), 3), "\n")
```

# High-Dimensional Data with Horseshoe Priors

```{r horseshoe-example}
# Generate high-dimensional data
set.seed(2025)
n_hd <- 300
p_hd <- 50  # Many covariates

# Create high-dimensional data frame
data_hd <- as.data.frame(matrix(rnorm(n_hd * p_hd), n_hd, p_hd))
names(data_hd) <- paste0("X", 1:p_hd)

# Only first few covariates are truly relevant
data_hd$A <- rbinom(n_hd, 1, plogis(
  1.0*data_hd$X1 + 1.0*data_hd$X2 - 1.0*data_hd$X3 + 1.0*data_hd$X4
))

# Outcome depends on only a few variables (sparse)
data_hd$Y <- 2 + 1.5*data_hd$A + 
             1.0*data_hd$X1 + 1.0*data_hd$X2 + 1.0*data_hd$X3 + 
             1.0*data_hd$A*data_hd$X1 + 
             rnorm(n_hd, 0, 1)

# Create explicit formulas for high-dimensional data
# Get all covariate names (excluding A and Y)
covariate_names <- setdiff(names(data_hd), c("A", "Y"))
outcome_formula_hd <- as.formula(paste("Y ~ A +", paste(covariate_names, collapse = " + ")))
ps_formula_hd <- as.formula(paste("A ~", paste(covariate_names, collapse = " + ")))

# Standard priors (might overfit)
result_standard <- DRBayes.PC(
  outcome.formula = outcome_formula_hd,  # Include all covariates
  ps.formula = ps_formula_hd,            # All covariates except A and Y
  data = data_hd,
  outcome.model = B.LM,
  ps.model = B.Logit,
  mc = 5000, bn = 1000, thin = 2
)

# Horseshoe priors for regularization
result_horseshoe <- DRBayes.PC(
  outcome.formula = outcome_formula_hd,
  ps.formula = ps_formula_hd,
  data = data_hd,
  outcome.model = HS.LM,        # Horseshoe linear regression
  ps.model = HS.Logit,          # Horseshoe logistic regression
  mc = 5000, bn = 1000, thin = 2
)

cat("High-dimensional results:\n")
cat("True ATE: 1.5\n")
cat("Standard priors ATE:", round(mean(result_standard$pc), 3), "\n")
cat("Horseshoe priors ATE:", round(mean(result_horseshoe$pc), 3), "\n")
cat("Standard priors SD:", round(sd(result_standard$pc), 3), "\n")
cat("Horseshoe priors SD:", round(sd(result_horseshoe$pc), 3), "\n")
```

# Results and Interpretation

## Summary Statistics

```{r results-summary}
# Create summary function
summarize_results <- function(samples, method_name, true_value = NULL, outcome_type = "continuous") {
  cat("\n", method_name, ":\n")
  cat("  Mean:", round(mean(samples), 3), "\n")
  cat("  SD:", round(sd(samples), 3), "\n")
  cat("  95% CI: [", round(quantile(samples, 0.025), 3), ",", 
      round(quantile(samples, 0.975), 3), "]\n")
  
  if (!is.null(true_value)) {
    bias <- mean(samples) - true_value
    cat("  Bias:", round(bias, 3), "\n")
    coverage <- (true_value >= quantile(samples, 0.025)) & 
                (true_value <= quantile(samples, 0.975))
    cat("  95% CI Coverage:", coverage, "\n")
  }

  if (outcome_type == "binary") {
    cat("  Interpretation: Treatment increases success probability by", 
        round(mean(samples), 3), "on average\n")
  }
}

# Summarize continuous outcome results
cat("=== CONTINUOUS OUTCOME RESULTS ===")
summarize_results(result_continuous$g.comp, "G-computation", true_ate_continuous)
summarize_results(result_continuous$pc, "Posterior Coupling", true_ate_continuous)

# Summarize binary outcome results  
cat("\n=== BINARY OUTCOME RESULTS ===")
summarize_results(result_binary_logit$g.comp, "G-computation (Logit)", true_ate_binary, "binary")
summarize_results(result_binary_logit$pc, "Posterior Coupling (Logit)", true_ate_binary, "binary")
summarize_results(result_binary_probit$pc, "Posterior Coupling (Probit)", true_ate_binary, "binary")
```

## Posterior Distributions

```{r posterior-plots, fig.width=12, fig.height=8}
# Create data frame for plotting
plot_data <- data.frame(
  ATE = c(result_continuous$g.comp, result_continuous$pc,
          result_binary_logit$g.comp, result_binary_logit$pc),
  Method = rep(c("G-computation", "Posterior Coupling"), 2),
  Outcome_Type = rep(c("Continuous", "Continuous", "Binary", "Binary"), 
                     each = length(result_continuous$g.comp))
)

# Create faceted density plot
p1 <- ggplot(plot_data, aes(x = ATE, fill = Method)) +
  geom_density(alpha = 0.7) +
  geom_vline(data = data.frame(
    Outcome_Type = c("Continuous", "Binary"),
    true_ate = c(true_ate_continuous, true_ate_binary)
  ), aes(xintercept = true_ate), linetype = "dashed", 
             color = "red", linewidth = 1) +
  facet_wrap(~ Outcome_Type, scales = "free") +
  labs(
    title = "Posterior Distributions of Average Treatment Effect",
    subtitle = "Red line shows true ATE",
    x = "Average Treatment Effect",
    y = "Density",
    fill = "Method"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

print(p1)
```

## Link Function Comparison (Binary Outcomes)

```{r link-comparison-plot, fig.width=10, fig.height=6}
# Compare logit vs probit for binary outcomes
link_data <- data.frame(
  ATE = c(result_binary_logit$pc, result_binary_probit$pc),
  Link = rep(c("Logit", "Probit"), each = length(result_binary_logit$pc))
)

p2 <- ggplot(link_data, aes(x = ATE, fill = Link)) +
  geom_density(alpha = 0.7) +
  geom_vline(xintercept = true_ate_binary, linetype = "dashed", 
             color = "red", linewidth = 1) +
  labs(
    title = "Comparison of Link Functions for Binary Outcomes",
    subtitle = "Red line shows true ATE (probability difference)",
    x = "Average Treatment Effect (Probability Difference)",
    y = "Density",
    fill = "Link Function"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

print(p2)
```

## Horseshoe vs Standard Priors

```{r horseshoe-comparison-plot, fig.width=10, fig.height=6}
# Compare standard vs horseshoe priors for high-dimensional data
horseshoe_data <- data.frame(
  ATE = c(result_standard$pc, result_horseshoe$pc),
  Prior = rep(c("Standard", "Horseshoe"), each = length(result_standard$pc))
)

p3 <- ggplot(horseshoe_data, aes(x = ATE, fill = Prior)) +
  geom_density(alpha = 0.7) +
  geom_vline(xintercept = 1.5, linetype = "dashed", 
             color = "red", linewidth = 1) +
  labs(
    title = "Standard vs Horseshoe Priors for High-Dimensional Data",
    subtitle = "Red line shows true ATE = 1.5",
    x = "Average Treatment Effect",
    y = "Density",
    fill = "Prior Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

print(p3)
```

## Convergence Diagnostics

```{r convergence-diagnostics, fig.width=12, fig.height=8}
# Create trace plots for both outcome types
par(mfrow = c(2, 2))

# Continuous outcome trace plots
plot(result_continuous$pc, type = "l", 
     main = "Trace Plot: Continuous Outcome (PC)",
     xlab = "Iteration", ylab = "ATE", col = "blue")
abline(h = true_ate_continuous, col = "red", lty = 2)

plot(result_continuous$g.comp, type = "l", 
     main = "Trace Plot: Continuous Outcome (G-comp)",
     xlab = "Iteration", ylab = "ATE", col = "blue")
abline(h = true_ate_continuous, col = "red", lty = 2)

# Binary outcome trace plots
plot(result_binary_logit$pc, type = "l", 
     main = "Trace Plot: Binary Outcome (PC)",
     xlab = "Iteration", ylab = "ATE (Prob. Diff.)", col = "blue")
abline(h = true_ate_binary, col = "red", lty = 2)

plot(result_binary_logit$g.comp, type = "l", 
     main = "Trace Plot: Binary Outcome (G-comp)",
     xlab = "Iteration", ylab = "ATE (Prob. Diff.)", col = "blue")
abline(h = true_ate_binary, col = "red", lty = 2)

par(mfrow = c(1, 1))
```

```{r convergence-stats}
# Basic convergence statistics
cat("Convergence Diagnostics:\n")
cat("Continuous Outcome:\n")
cat("  PC effective sample size:", 
    round(length(result_continuous$pc) / (1 + 2 * sum(acf(result_continuous$pc, plot = FALSE)$acf[-1][1:min(10, length(result_continuous$pc)-1)]))), "\n")
cat("  G-comp effective sample size:", 
    round(length(result_continuous$g.comp) / (1 + 2 * sum(acf(result_continuous$g.comp, plot = FALSE)$acf[-1][1:min(10, length(result_continuous$g.comp)-1)]))), "\n")

cat("Binary Outcome:\n")
cat("  PC effective sample size:", 
    round(length(result_binary_logit$pc) / (1 + 2 * sum(acf(result_binary_logit$pc, plot = FALSE)$acf[-1][1:min(10, length(result_binary_logit$pc)-1)]))), "\n")
cat("  G-comp effective sample size:", 
    round(length(result_binary_logit$g.comp) / (1 + 2 * sum(acf(result_binary_logit$g.comp, plot = FALSE)$acf[-1][1:min(10, length(result_binary_logit$g.comp)-1)]))), "\n")
```

# Model Comparison

## Comparison with Classical Methods

```{r classical-comparison}
# ===== CONTINUOUS OUTCOME COMPARISON =====
# Simple regression adjustment
lm_model <- lm(Y ~ A + X1 + X2 + X3 + A:X1, data = data)
reg_adj_ate_cont <- coef(lm_model)["A"]

# Inverse probability weighting
ps_model <- glm(A ~ X1 + X2 + X3, family = binomial, data = data)
ps_hat <- predict(ps_model, type = "response")
ipw_ate_cont <- mean(data$Y * data$A / ps_hat) - mean(data$Y * (1-data$A) / (1 - ps_hat))

# ===== BINARY OUTCOME COMPARISON =====
# Logistic regression adjustment
glm_model <- glm(Y ~ A + X1 + X2 + X3, family = binomial, data = data_binary)
newdata_1 <- newdata_0 <- data_binary[c("A", "X1", "X2", "X3")]
newdata_1$A <- 1
newdata_0$A <- 0
prob1_hat <- predict(glm_model, newdata = newdata_1, type = "response")
prob0_hat <- predict(glm_model, newdata = newdata_0, type = "response")
reg_adj_ate_bin <- mean(prob1_hat - prob0_hat)

# IPW for binary outcome (using same PS model structure)
ps_model_bin <- glm(A ~ X1 + X2 + X3, family = binomial, data = data_binary)
ps_hat_bin <- predict(ps_model_bin, type = "response")
ipw_ate_bin <- mean(data_binary$Y * data_binary$A / ps_hat_bin) - 
               mean(data_binary$Y * (1-data_binary$A) / (1 - ps_hat_bin))

# Create comparison tables
comparison_continuous <- data.frame(
  Method = c("True ATE", "Regression Adjustment", "IPW", 
             "DR (Bayesian, G-comp)", "DR (Bayesian, PC)"),
  Estimate = c(true_ate_continuous, reg_adj_ate_cont, ipw_ate_cont,
               mean(result_continuous$g.comp), mean(result_continuous$pc)),
  SE_or_SD = c(NA, summary(lm_model)$coefficients["A", "Std. Error"], 
               NA, sd(result_continuous$g.comp), sd(result_continuous$pc))
)

comparison_binary <- data.frame(
  Method = c("True ATE", "Regression Adjustment", "IPW",
             "DR (Bayesian, G-comp)", "DR (Bayesian, PC)"),
  Estimate = c(true_ate_binary, reg_adj_ate_bin, ipw_ate_bin,
               mean(result_binary_logit$g.comp), mean(result_binary_logit$pc)),
  SE_or_SD = c(NA, NA, NA, 
               sd(result_binary_logit$g.comp), sd(result_binary_logit$pc))
)

# Round numeric columns
comparison_continuous[, c("Estimate", "SE_or_SD")] <- round(comparison_continuous[, c("Estimate", "SE_or_SD")], 3)
comparison_binary[, c("Estimate", "SE_or_SD")] <- round(comparison_binary[, c("Estimate", "SE_or_SD")], 3)

cat("=== CONTINUOUS OUTCOME COMPARISON ===\n")
print(comparison_continuous)

cat("\n=== BINARY OUTCOME COMPARISON ===\n")  
print(comparison_binary)
```

# Practical Considerations

## Formula Specification Guidelines

Writing formulas in DRBayes is like telling the software how you think your data works. Here's a practical guide to help you build effective models.

### Step-by-Step Approach to Building Your Outcome Model

**Step 1: Start Simple**
```r
# Always include the treatment variable A - this is essential!
outcome.formula = Y ~ A + X1 + X2 + X3
```
- **What this does**: Models Y as depending on treatment A plus your main confounders
- **When to use**: This should be your starting point for any analysis
- **Think of it as**: "My outcome depends on treatment plus the main factors I care about"

**Step 2: Add Treatment Interactions (if relevant)**
```r
# Add interactions when treatment effects might vary by patient characteristics
outcome.formula = Y ~ A + X1 + X2 + X3 + A:X1
```
- **What this does**: Allows the treatment effect to be different for different values of X1
- **When to use**: When you suspect treatment works better/worse for certain groups
- **Real example**: "Blood pressure medication might work differently for older vs younger patients"
- **Interpretation**: If A:X1 coefficient is positive, treatment is more effective when X1 is higher

**Step 3: Handle Non-Linear Relationships (when needed)**
```r
# Use transformations when relationships aren't straight lines
outcome.formula = Y ~ A + log(X1) + I(X2^2) + X3 + A:X1

# Or use polynomials for smooth curves
outcome.formula = Y ~ A + X1 + poly(X2, 2) + X3 + A:X1
```
- **log(X1)**: Use when X1 has diminishing returns (e.g., income effects)
- **I(X2^2)**: Use when you expect U-shaped or inverted-U relationships (e.g., age effects)
- **poly(X2, 2)**: Similar to X2^2 but mathematically more stable
- **Warning**: Don't use log() with zero or negative values!

**Step 4: Advanced Specifications (use carefully)**
```r
# Only when you have strong reasons and enough data
outcome.formula = Y ~ A + X1 + X2 + X3 + A:X1 + A:X2 + I(X1^2) + X1:X2
```
- **When to use**: Large datasets (n > 500) with clear theoretical justification
- **Risk**: Overfitting - your model might work well on your data but poorly on new data

### Building Your Propensity Score Model

The propensity score model predicts who gets treatment. Think: "What factors influenced the treatment decision?"

**Start Here:**
```r
# Include all variables that affect BOTH treatment assignment AND outcome
ps.formula = A ~ X1 + X2 + X3
```
- **Key principle**: Include confounders (variables affecting both treatment and outcome)
- **Don't include**: Pure predictors of outcome that don't affect treatment assignment
- **Don't include**: The outcome variable Y itself

**Add Complexity When Needed:**
```r
# Non-linear relationships in treatment assignment
ps.formula = A ~ X1 + X2 + X3 + I(X1^2) + X1:X2
```
- **When to use**: When treatment assignment has complex patterns
- **Example**: "Doctors might prescribe more aggressively for moderate-risk patients but less for very high-risk patients"

**High-Dimensional Data:**
```r
# Include many variables (use with horseshoe priors)
ps.formula = A ~ .  # All variables except A and Y
```
- **When to use**: Rich datasets with many potential confounders
- **Requirement**: Must use regularized priors (HS.Logit, HS.Probit)
- **Example**: Electronic health records with hundreds of variables

### Quick Decision Guide

**"Should I include this variable in my model?"**

For Outcome Model:
- ✅ **Always include**: Treatment variable A
- ✅ **Usually include**: Strong predictors of outcome
- ✅ **Consider including**: Variables that modify treatment effects
- ❌ **Don't include**: Variables measured after treatment

For Propensity Score Model:
- ✅ **Always include**: Confounders (affect both treatment and outcome)
- ✅ **Usually include**: Variables that predict treatment assignment
- ❌ **Never include**: The outcome variable itself

**"Should I add interactions/transformations?"**
- ✅ **Yes, if**: You have theoretical reasons and sufficient sample size
- ✅ **Yes, if**: Diagnostic plots suggest non-linear relationships
- ❌ **No, if**: Your sample size is small (n < 200)
- ❌ **No, if**: You're just ``fishing'' without theoretical justification

### Common Mistakes to Avoid

1. **Forgetting the treatment variable**: Always include A in your outcome formula
2. **Including post-treatment variables**: Don't include variables measured after treatment
3. **Over-complicating with small samples**: Keep it simple with n < 500
4. **Ignoring domain knowledge**: Use subject-matter expertise, not just statistical significance

## MCMC Recommendations

- **MCMC iterations**: At least 2000 after burn-in for stable inference
- **Burn-in**: Typically 20-25% of total iterations (e.g., 1000 out of 4000)
- **Thinning**: Use `thin = 2` or `thin = 3` to reduce autocorrelation
- **High-dimensional data**: May require more iterations and stronger priors

## Missing Data Considerations

```{r missing-data-options}
# Different missing data strategies
cat("Missing data handling options:\n")
cat('na.action = "na.omit"     # Remove rows with any missing values (default)\n')
cat('na.action = "na.fail"     # Stop if any missing values found\n')
cat('na.action = "na.exclude"  # Similar to na.omit but preserves attributes\n')

# Check data completeness before analysis
cat("\nData completeness check:\n")
cat("Complete cases in original data:", sum(complete.cases(data)), "out of", nrow(data), "\n")
cat("Complete cases in missing data:", sum(complete.cases(data_missing)), "out of", nrow(data_missing), "\n")
```

## Troubleshooting Common Issues

```{r diagnostics}
# Check for extreme values
cat("Posterior diagnostics:\n")
cat("Continuous outcome range:", round(range(result_continuous$pc), 3), "\n")
cat("Binary outcome range:", round(range(result_binary_logit$pc), 3), "\n")

# Check for reasonable credible intervals
continuous_ci <- quantile(result_continuous$pc, c(0.025, 0.975))
binary_ci <- quantile(result_binary_logit$pc, c(0.025, 0.975))

cat("95% CI width (continuous):", round(diff(continuous_ci), 3), "\n")
cat("95% CI width (binary):", round(diff(binary_ci), 3), "\n")

# Basic autocorrelation check
acf_cont <- acf(result_continuous$pc, plot = FALSE, lag.max = 5)
acf_bin <- acf(result_binary_logit$pc, plot = FALSE, lag.max = 5)

cat("Lag-1 autocorrelation (continuous):", round(acf_cont$acf[2], 3), "\n")
cat("Lag-1 autocorrelation (binary):", round(acf_bin$acf[2], 3), "\n")

# Effective sample size estimation
eff_cont <- length(result_continuous$pc) / (1 + 2 * sum(acf_cont$acf[-1]))
eff_bin <- length(result_binary_logit$pc) / (1 + 2 * sum(acf_bin$acf[-1]))

cat("Effective sample size (continuous):", round(max(eff_cont, 1)), "\n")
cat("Effective sample size (binary):", round(max(eff_bin, 1)), "\n")
```

# Alternative: Bayesian Bootstrap Method

```{r bayesian-bootstrap}
# Compare with the Bayesian Bootstrap approach
# Prepare data in the format expected by DRBayes.BB
X.lm_bb <- data.frame(
  A = data$A,
  X1 = data$X1, X2 = data$X2, X3 = data$X3,
  A_X1 = data$A * data$X1  # Interaction term
)
X.ps_bb <- data.frame(X1 = data$X1, X2 = data$X2, X3 = data$X3)

result_bb <- DRBayes.BB(
  Y = data$Y,
  A = data$A,
  X.lm = X.lm_bb,
  X.ps = X.ps_bb,
  num_iterations = 2000,
  family = "gaussian"
)

cat("Bayesian Bootstrap ATE:", round(mean(result_bb), 3), "\n")
cat("Bayesian Bootstrap SD:", round(sd(result_bb), 3), "\n")
cat("95% CI: [", round(quantile(result_bb, c(0.025, 0.975)), 3), "]\n")

# Compare all methods
cat("\nMethod Comparison (Continuous Outcome):\n")
cat("True ATE:", round(true_ate_continuous, 3), "\n")
cat("Posterior Coupling:", round(mean(result_continuous$pc), 3), "\n")
cat("G-computation:", round(mean(result_continuous$g.comp), 3), "\n")
cat("Bayesian Bootstrap:", round(mean(result_bb), 3), "\n")
```

# Conclusion

The DRBayes package with its formula interface provides a powerful and intuitive framework for Bayesian doubly robust causal inference. Key advantages include:

## Formula Interface Benefits
- **Familiar syntax**: Compatible with `lm()`, `glm()`, and other R functions
- **Automatic handling**: Design matrices, factor variables, interactions, transformations
- **Missing data support**: Built-in handling with flexible `na.action` options
- **Complex modeling**: Easy specification of non-linear relationships

## Methodological Advantages
- **Robustness**: Consistent estimation under model misspecification
- **Uncertainty quantification**: Full posterior distributions for treatment effects  
- **Outcome flexibility**: Support for both continuous and binary outcomes with multiple link functions
- **Regularization**: Horseshoe priors for high-dimensional data
- **Software integration**: Works with built-in models or external Bayesian software

## Practical Recommendations
1. **Start simple**: Use basic formulas and automatic detection
2. **Check diagnostics**: Monitor convergence and effective sample sizes
3. **Use appropriate priors**: Horseshoe priors for high-dimensional data
4. **Handle missing data**: Choose appropriate `na.action` strategy
5. **Compare methods**: Use both posterior coupling and g-computation for robustness

For integration with external Bayesian software (Stan, JAGS, brms), see the additional vignettes.

# Session Information

```{r session-info}
sessionInfo()
```
